Untitled CAV Paper

### 2016-01-21 paper.tex
Beginning of the section of unbounded synthesis

Advances in SAT solving technology has prompted research into its applicability
to synthesis as an alternative to BDDs. One approach is to find sets of states
in CNF \cite{demiurge}. Another approach is to eschew states and focus on
\emph{runs} of the game. Previous work has applied this idea to realizability
of bounded games \cite{nina} by forming abstract representations of the game.
In this paper, we extend this idea to unbounded games by constructing
approximate sets of winning states from abstract trees.

@@

A \emph{safety game}, $G = \langle X, L_c, L_u, \delta, I, E, \rangle$,
consists of a set of state variables, sets of controllable and uncontrollable
label variables, a transition relationship $\delta : (X, L_c, L_u) \to X$, an
initial state, and an error state. The \emph{controller} and \emph{environment}
players choose controllable and uncontrollable labels respectively and the game
proceeds according to $\delta$.

@@

An abstract game tree can be checked for the existence of a winning run by a
SAT solver.  The tree must be encoded into CNF by making copies of the
transition relation for each game step in tree. The formula must also check
whether the error state has been reached in no branch/in all branches for the
controller and environment respectively. The functions that produce these
formulas are shown in Algorithm \ref{alg:treeFormula}.

@@

Synthesis begins with an empty game as seen in Figure \ref{fig:emptytree}.
Initially we are playing on behalf of the environment because it chooses the
first move in each step. The empty game is passed to the SAT solver, which
searches for a candidate environment strategy.  If a candidate is found (Figure
\ref{fig:candidate}) then it is checked for a spoiling strategy by solving for
the controller. If no spoiling strategy exists, that means our candidate is a
winning strategy and the algorithm terminates.  Otherwise we find a
counterexample (Figure \ref{fig:cex}), which is used to refine the empty game
tree to include the first move from the controller's spoiling strategy (Figure
\ref{fig:refined}.  The algorithm continues by finding a new candidate for the
environment (Figure \ref{fig:candidate2}), and a new counterexample to refine
the game abstraction once again (Figure \ref{fig:refined2}).

@@

>>>>

Advances in SAT solving technology has prompted research into its applicability
to synthesis as an alternative to BDDs. One approach is to find sets of states
in CNF \cite{demiurge}. Another approach is to eschew states and focus on
\emph{runs} of the game. Previous work has applied this idea to realizability
of bounded games \cite{narodytska2014} by forming abstract representations of
the game.  In this paper, we extend this idea to unbounded games by
constructing approximate sets of winning states from abstract trees.

@@

A \emph{safety game}, $G = \langle X, L_c, L_u, \delta, I, E, \rangle$,
consists of a set of boolean state variables, sets of controllable and
uncontrollable label variables, a transition relationship $\delta : (X, L_c,
L_u) \to X$, an initial state, and an error state. The \emph{controller} and
\emph{environment} players choose controllable and uncontrollable labels
respectively and the game proceeds according to $\delta$.

@@

An abstract game tree can be checked for the existence of a winning run by a
SAT solver\cite{narodytska2014}.  The tree must be encoded into CNF by making
copies of the transition relation for each game step in tree. The formula must
also check whether the error state has been reached in no branch/in all
branches for the controller and environment respectively. The functions that
produce these formulas are shown in Algorithm \ref{alg:treeFormula}.

@@

The bounded synthesis algorithm described in \cite{narodytska2014} begins with
an empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing
on behalf of the environment because it chooses the first move in each step.
The empty game is passed to the SAT solver, which searches for a candidate
environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
then it is checked for a spoiling strategy by solving for the controller.  If
no spoiling strategy exists, that means our candidate is a winning strategy and
the algorithm terminates.  Otherwise we find a counterexample (Figure
\ref{fig:cex}), which is used to refine the empty game tree to include the
first move from the controller's spoiling strategy (Figure \ref{fig:refined}.
The algorithm continues by finding a new candidate for the environment (Figure
\ref{fig:candidate2}), and a new counterexample to refine the game abstraction
once again (Figure \ref{fig:refined2}).

@@

\section{Unbounded Synthesis}

Bounded synthesis can be used to prove the existence of a winning strategy for
the environment by providing a witness. For the controller, the strongest claim
that can be made is that the strategy is winning as long as the game does not
extend longer than the bound.

In model checking the maximum bound is decided based on the states of the game
\cite{biere1999}. The na\"ive approach is to use size of the state space as the
bound ($2^{|X|}$). With a bound of this size all states may be explored. One
optimisation is to use the diameter of the game, which is the smallest number
$d$ such that for any state $x$ there is a path of length $\leq d$ to all other
reachable states. However, for large games these bounds are intractable.

When performing model checking or synthesis with BDDs \cite{burch1990} the set
of states that are winning for the environment is iteratively constructed by
computing the states from which the environment can force the game into the
previous winning set. Eventually this process reaches a fixed point and the
total set of environment winning states is known.

A similar concept can be applied to the bounded synthesis algorithm to
iteratively increase the bound of the game and terminate when a fixed point is
reached. When a strategy is found to be winning on an abstract game tree, we
record as winning the states from which the opponent could find no
counterexample. To find these states we use interpolation of subformulas of the
game tree.

\subsection{Learning States with Interpolants}

Given two formulas $A$ and $B$ such that $A \land B$ is unsatisfiable, it is
possible to construct and interpolant $\mathcal{I}$ such that $A \to
\mathcal{I}$, $B \land \mathcal{I}$ is unsatisfiable, and $\mathcal{I}$ refers
only to the intersection of variables in $A$ and $B$. An interpolant can be
constructed efficiently from a resolution proof of the unsatisfiability of $A
\land B$ \cite{pudlak1997}.

\bibliographystyle{splncs03}
\bibliography{paper}
